\ppse{} - A Peer-to-Peer Support for Multiplayer Games

### 2008-09-01 textos/p2pse-sbgames2008/ppsg.tex

The first layer in the \ppse{} architecture is responsible for forming the mesh of point-to-point connections. The client's connections with the servers and with other clients are established and managed in this layer. The basic network functions for connection establishment, such as socket creation, connection and packet sending and receiving, are given by a low-level API provided by the ZIG library \cite{zige} which basically implements a protocol that resembles SCTP \cite{stewart2000sct} over UDP.

The Layer 0 server is responsible for managing connections with all clients. Naturally, it maintains a list of all clients that are connected to it along with a logical communication endpoint object for each client. When started, the server spawns a socialization space called the social space object. All clients that are not in an action space are bound to the social space. All communication between clients that are in the social space is intermediated by the server.

@@

\begin{itemize}
\item The server sends a syn message to the new member and to all the other members informing them to which clients they must be connected in order to enter (or stay) in the group. The server sends a list with all the members of the group;
\item All members then try to establish connections to every other member in the recently received member list, or just keep the connection, if it was already established (the ZIG library supports simultaneous bidirectional connections through the same socket);
\item When a client is entering a group, but still has not finished the process, it belongs to the ``limbo'';
\item As soon as the new member establishes connections to the other clients, the server excludes it from the limbo and inserts in the member list of the group;
\item When a client is removed from the group, all connections to the other clients are interrupted and it returns to the social space;
\item If a client, for any reason, loses its connection to the server, it is removed from the group;
\item When a client enters or leaves a group, the server sends to the others  an up-to-date member list.
\end{itemize}

@@

Layer 0 is also responsible for the timeout control of the connections. Keep-alive dummy messages are sent regularly in order to maintain all connections active. If, after an interval of time specified by the application, no messages from a host have arrived, the connection is closed.

@@

Depending on the game, the resulting conflicts from a lack of centralized timestamping of movement and similar events, whenever they ensue, can be either handled by each peer individually or left for the super-peer to detect it and issue a special correction message later, as the BZFlag protocol does \cite{pellegrino2003bra}. Local corrections are an option if each peer is authoritative over its own avatar's position and if it broadcasts position updates for its own avatar. For instance, if a peer's local avatar overlaps with a remote peer's, the local avatar can push its own avatar out of that position. The remote avatar will do the same. The only issue is ensuring that the correction step won't make both peers try to resolve the situation by unstucking their avatars to the same spot again. This local conflict resolution is what we are currently employing in our capture-the-flag action prototype, \hoverkill\ (reference blinded).%Hoverkill \cite{hoverkill}.

@@

In the previous section, we revealed that we encourage peers in an action space to send authoritative position updates, among other events that migh! have low-latency requirements and a low probability of having their order of execution affecting the resulting game states too much (weakly-consistent events). This opens up the possibility of cheating.

In this project we tried a somewhat unique approach [reference blinded]%\cite{gaspareto2008nna}
. Since most events would be still centralized at the super-peer and most action games have one-per-client avatar position updates as their consistency, bandwidth and latency bottlenecks, we just left peers with total control over their positions and with the responsibility for sending them directly to other peers in a broadcast fashion. This optimizes for consistency and latency, and doesn't create a bandwidth bottleneck at the super-peer. On top of that, we tried AI-based detection of unusual patterns over streams of position updates coming from each peer. The approach used for cheat detection was focused on the data packets being transferred during the game sessions - both in the transmitters and receptors - and how they are used during the functioning of the system. Only packets relevant to the cheat being investigated are considered.

In this manner, the cheat detection module is implemented in an independent way and works as an attachment to \ppse{}, which, with a parameter passage protocol, informs the set of data received from a player and returns a correspondent ``fraud index'', that tries to quantify the cheating that the player is assumed to have performed in the analyzed period.

The adopted approach was to consider only movement data as input to the cheat detection module. Although this might seem to be too few information, it is based on movement that  malicious players perform great part of their cheating. For example, one could try to move at a speed higher than the maximum speed allowed in the game (most common kind of cheating in MMOGs) or teleport instantaneously to strategic spots, also known as speed-cheating.

This way, the point is to analyze a set of movements of the player, in order to detect anomalies in his moves. Each player will be evaluated individually, based on his position history, using Artificial Neural Networks (ANN). ANNs, because of their great generalization and pattern recognition capacity, present satisfactory results in tasks similar to the speed-cheating problem. The idea is to identify, in the set of analyzed data, at least one movement that could be considered a fraud, which characterizes the player where this set of data came from as a cheater. The ANN then identifies and classifies the movement patterns as acceptable and non-acceptable.

During the modeling and design of the cheat detection module, the architectures of the ANNs, the set of training data and tests have been defined in order to allows them to solve this kind of problem. Considering that the task consists of the classification of behavior patterns, it has been decided to use the full-connected multilayer perceptron ANN with supervised learning based on back-propagation. The reason to choose it was that it is the ANN model most commonly used in many classification applications, due to it simplicity of understanding and implementation.

@@

The main purpose of the security module is to provide secure communications between server and clients, and between peer clients. Our security module is almost completely transparent to the application, both in terms of API and in network and CPU overhead. The module does many tasks behind the scenes, such as cryptographic key distribution and secure channel handshakes, tending to each application's individual needs for message confidentiality, integrity and authenticity.

@@

The module has been implemented by extending the ZIG library functionalities, providing secure communication, transparently to the application, being responsible for distributing cryptogaphic keys. It runs over HawkNL \cite{hawknl} and over SEC\ppse{}. HawkNL is a library that, among other functionalities, provides a simpler and portable API for sockets. SEC\ppse{} is our library that provides C++ abstractions for cryptographic algorithms available in Nettle \cite{nettle} and CyoEncode \cite{cyotec}. SEC\ppse{} also implements certification and serialization functionalities meant for security related transmissions. Figure \ref{fig:security} illustrates the security module architecture.%Nettle is a low-level cryptographic library written in C, which offers algorithms for hashing (MD5, SHA1, SHA256 etc.), symmetric (AES, ARCFOUR, 3DES etc.) and assymetric (HMAC-MD5, HMAC-SHA-1 and HMAC-SHA-256) cryptography and generation of pseudo-random data (Yarrow). CyoEncode \cite{cyotec} is a small library which provides coding and decoding in Base64, Base32 and Base16.

In our secure communication protocol, all packets sent, both by clients and by servers, are intercepted right before being transmitted through the network (packets generated by ZIG). In each one of them, a small header is introduced with control information used to guarantee confidentiality, integrity and authenticity of data when such attributes are required. The size of this header is variable, for it depends on what characteristics are desired by the application, and on the intended level of security (key length). For a real-time application, such as distributed games, usually there is no need for a very long key, because the hacker would have too little time to break the algorithm by brute force. For example, using AES and HMAC-MD5, it would be necessary only 21 bytes. In general, the overhead size of each packet is somewhere between 1 and 37 bytes. These packets containing user data and control header are then called \textbf{data packets}.

@@

Every time that a peer needs to communicate with another which was unknown until that time, they first need to negotiate a key. Therefore, instead of sending the data packet required by the application, the library sends a requisition for the certificate of the destination through a control packet, which already attaches the source's own certificate. When the packet arrives in the destination peer, the security library reads the attached certificate and verifies its authenticity with  the specified certification authority responsible for the game. If the verification fails, the packet is discarded. If it succeeds, the process continues: the peer who received the packet allocates a reply packet, where it attaches its own certificate, as required, and a randomly generated key, which is encrypted using the public key present in the certificate of the peer that initiated the whole process. This way, it can be assured that only the receiver of this reply packet will be able to decrypt it and then obtain the key. To guarantee that the sender of this packet is not an attacker who intercepted the communication, the already encrypted key is re-encrypted, this time with the private key of the sender. This way, both authenticity and integrity are provided.

After receiving the reply packet containing the double-encrypted key, the first peer decrypts it using its private key and then decrypts again, using the public key of the other peer, verifying its authenticity. When this whole process is successfully completed, the peers will be able to securely communicate with a fast symmetric cryptographic algorithm using the generated key.

@@

Since one of the main objectives of the \ppse{} engine is to reduce the bandwidth utilization by the server, a numerical comparison between the load on a traditional server and the load on a P2PSE server was required. In order to do that, it was necessary to simulate two game servers, each one with players connected to it. One of them was a traditional game server, to which the clients kept connected and trading packets the whole time. The other one was a \ppse{} server, where the players needed to trade packets with the server only while they were in the social space. In the action space, the players formed a peer-to-peer group, communicating to one another directly.

>>>>

The first layer in the \ppse{} architecture is responsible for forming the mesh of point-to-point connections. The clients' connections with the servers and with other clients are established and managed in this layer. The basic network functions for connection establishment, such as socket creation, connection and packet sending and receiving, are given by a low-level API provided by the ZIG library \cite{zige} which basically implements a protocol that resembles SCTP \cite{stewart2000sct} over UDP.

The Layer 0 server is responsible for managing connections with all clients. Naturally, it maintains a list of all clients that are connected to it along with a logical communication endpoint object for each client. When started, the server spawns a socialization space called the social space object. The clients that are not in an action space are bound to the social space - or to the ``limbo'', when in transit from one space to the other. All communication between clients that are in the social space is intermediated by the server.

@@

\begin{enumerate}
\item The server sends a synchronization message to the new member and to all the other members informing them to which clients they must be connected in order to enter (or stay) in the group. The server sends a list with all the members of the group;
\item All members then try to establish connections to every other member in the recently received member list, or just keep the connection, if it has already been established (the ZIG library supports simultaneous bidirectional connections through the same socket);
\item When a client is entering a group, but still has not finished the process, it belongs to the ``limbo'';
\item As soon as the new member establishes connections to the other clients, the server excludes it from the limbo and inserts it in the member list of the group;
\item When a client is removed from the group, all connections to the other clients are interrupted and it returns to the social space;
\item If a client, for any reason, loses its connection to the server, it is removed from the group;
\item When a client enters or leaves a group, the server sends to the others  an up-to-date member list.
\end{enumerate}

@@

Layer 0 is also responsible for the timeout control of the connections. Keep-alive dummy messages are sent regularly in order to maintain all connections active. If, after an interval of time specified by the application, no messages from a host have arrived, the connection to that host is closed.

@@

Depending on the game, the resulting conflicts from a lack of centralized timestamping of movement and similar events, whenever they ensue, can be either handled by each peer individually or left for the super-peer to detect it and issue a special correction message later, as the BZFlag protocol does \cite{pellegrino2003bra}. Local corrections are an option if each peer is authoritative over its own avatar's position and if it broadcasts position updates for its own avatar. For instance, if a peer's local avatar overlaps with a remote peer's, the local avatar can push its own avatar out of that position. The remote avatar will do the same. The only issue is ensuring that the correction step won't make both peers try to resolve the situation by unstucking their avatars to the same spot again. This local conflict resolution is what we are currently employing in our capture-the-flag action prototype, \hoverkill\ [reference blinded].%Hoverkill \cite{hoverkill}.

@@

In the previous section, we revealed that we encourage peers in an action space to send authoritative position updates, among other events that might have low-latency requirements and a low probability of having their order of execution affecting the resulting game states too much (weakly-consistent events). This opens up the possibility of cheating.

In this project we tried a somewhat unique approach [reference blinded]%\cite{gaspareto2008nna}
. Since most events would be still centralized at the super-peer and most action games have one-per-client avatar position updates as their latency, consistency and bandwidth bottlenecks, we just left peers with total control over their positions and with the responsibility for sending them directly to other peers in a broadcast fashion. This optimizes for consistency and latency, and doesn't create a bandwidth bottleneck at the super-peer. On top of that, we tried AI-based detection of unusual patterns over streams of position updates coming from each peer. The approach used for cheat detection was focused on the data packets being transferred during the game sessions - both in the transmitters and receptors - and how they are used during the functioning of the system. Only packets relevant to the cheat being investigated are considered.

The cheat detection module is implemented in an independent way and works as an attachment to \ppse{}, which, with a parameter passage protocol, informs the set of data received from a player and returns a correspondent ``fraud index'', that tries to quantify the cheating that the player is assumed to have performed in the analyzed period.

The adopted approach was to consider only movement data as input to the cheat detection module. Although this might seem to be too few information, it is based on movement that  malicious players perform great part of their cheating. For example, one could try to move at a speed higher than the maximum speed allowed in the game (most common kind of cheating in MMOGs) or teleport instantaneously to strategic spots, which is also known as speed-cheating.

This way, the point is to analyze a set of movements of the player, in order to detect anomalous sequences. Each player will be evaluated individually, based on his position history, using Artificial Neural Networks (ANN). ANNs, because of their great generalization and pattern recognition capacity, present satisfactory results in tasks similar to the speed-cheating problem. The idea is to identify, in the set of analyzed data, at least one movement that could be considered a fraud, which characterizes the player where this set of data came from as a cheater. The ANN then identifies and classifies the movement patterns as acceptable and non-acceptable.

During the modeling and design of the cheat detection module, the architectures of the ANNs, the set of training data and the set of tests have been defined in order to allows them to solve this kind of problem. Considering that the task consists of the classification of behavior patterns, it has been decided to use the full-connected multilayer perceptron ANN with supervised learning based on back-propagation. The reason to choose it was that it is the ANN model most commonly used in many classification applications, due to it simplicity of understanding and implementation.

@@

The main purpose of the security module is to provide secure communication between servers and clients, and between peers in an action space group. Our security module is almost completely transparent to the application, both in terms of API and in network and CPU overhead. The module does many tasks behind the scenes, such as cryptographic key distribution and secure channel handshakes, tending to each application's individual needs for message confidentiality, integrity and authenticity.

@@

The module has been implemented by extending the ZIG library functionalities, providing secure communication, transparently to the application, being responsible for distributing cryptogaphic keys. It runs over HawkNL \cite{hawknl} and over SEC\ppse{}. HawkNL is a library that, among other functionalities, provides a simpler and portable API for sockets. SEC\ppse{} is our library that provides C++ abstractions for cryptographic algorithms available in Nettle \cite{nettle} and \mbox{CyoEncode} \cite{cyoencode}. SEC\ppse{} also implements certification and serialization functionalities meant for security related transmissions. Figure \ref{fig:security} illustrates the security module architecture.%Nettle is a low-level cryptographic library written in C, which offers algorithms for hashing (MD5, SHA1, SHA256 etc.), symmetric (AES, ARCFOUR, 3DES etc.) and assymetric (HMAC-MD5, HMAC-SHA-1 and HMAC-SHA-256) cryptography and generation of pseudo-random data (Yarrow). CyoEncode \cite{cyotec} is a small library which provides coding and decoding in Base64, Base32 and Base16.

In our secure communication protocol, all ZIG-generated packets sent, both by clients and by servers, are intercepted right before being transmitted through the network. In each one of them, a small header is introduced with control information used to guarantee confidentiality, integrity and authenticity of data when such attributes are required. The size of this header is variable, for it depends on what characteristics are desired by the application, and on the intended level of security (key length). For a real-time application, such as distributed games, usually there is no need for a very long key, because the hacker would have too little time to break the algorithm by brute force. For example, using AES and HMAC-MD5, it would be necessary only 21 bytes. In general, the overhead size of each packet is somewhere between 1 and 37 bytes. These packets containing user data and control header are then called \textbf{data packets}.

@@

Every time that a peer needs to communicate with another which was unknown until that time, they first need to negotiate a key. Therefore, before sending the data packet required by the application, the library sends a requisition for the certificate of the destination through a control packet, which already carries the source's own certificate. When the packet arrives in the destination peer, the security library reads the attached certificate and verifies its authenticity with  the specified certification authority responsible for the game. If the verification fails, the packet is discarded. If it succeeds, the process continues: the peer who received the packet allocates a reply packet, where it attaches its own certificate, as required, and a randomly generated key, which is encrypted using the public key present in the certificate of the peer that initiated the whole process. This way, it can be assured that only the receiver of this reply packet will be able to decrypt it and then obtain the key. To guarantee that the sender of this packet is not an attacker who intercepted the communication, the already encrypted key is re-encrypted, this time with the private key of the sender. This way, both authenticity and integrity are provided.

After receiving the reply packet containing the double-encrypted key, the first peer decrypts it using the public key of the other peer, verifying the authenticity of the message, and then decrypts it again, using  its own private key. When this whole process is successfully completed, the peers will be able to securely communicate with a fast symmetric cryptographic algorithm using the generated key.

@@

Since one of the main objectives of the \ppse{} engine is to reduce the bandwidth utilization by the server, a numerical comparison between the load on a traditional server and the load on a \ppse{} server was required. In order to do that, it was necessary to simulate two game servers, each one with players connected to it. One of them was a traditional game server, to which the clients kept connected and trading packets the whole time. The other one was a \ppse{} server, where the players needed to trade packets with the server only while they were in the social space. In the action space, the players formed a peer-to-peer group, communicating to one another directly.
